Summary of the things that I learned

### 2024-11-15 200-category-theory.tex
Add a remark about univalence of L0, and string the remarks about univalence of the category of retracts together

\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_univalence}]
  If $ \overline \C $ is univalent, $ \C $ is univalent as well.
\end{lemma}
\begin{proof}
  The fully faithful embedding $ \iota_\C : \C \hookrightarrow \overline \C $ induces equivalences $ (X \cong Y) \simeq (\iota_\C(X) \cong \iota_\C(Y)) $. We also have an equivalence $ (X = Y) \simeq (\iota_\C(X) = \iota_\C(Y)) $, because, as it turns out, any equality between $ X $ and $ Y $ also preserves the identity of $ X $. Therefore, if $ \overline \C $ is univalent, we have a chain of equivalences
  \[ (X = Y) \simeq (\iota_\C(X) = \iota_\C(Y)) \simeq (\iota_\C(X) \cong \iota_\C(Y)) \simeq (X \cong Y), \]
  which shows that $ \C $ is univalent as well.
\end{proof}

\begin{remark}
  Note that the converse does not necessarily hold. Consider the commutative monoid consisting of the three matrices
  \[
    a = \begin{pmatrix}
      1 & 0\\0 & 1
    \end{pmatrix}, \quad
    b = \begin{pmatrix}
      1 & 0\\0 & 0
    \end{pmatrix} \quad \text{and} \quad
    c = \begin{pmatrix}
      -1 & 0\\0 & 0
    \end{pmatrix}
  \]
  under matrix multiplication. As we saw in Section \ref{sec:monoid-category}, we can turn this into a category $ \C_M $ with one object $ \star $, and the three morphisms $ a $, $ b $ and $ c $. Only $ a $ is an isomorphism and since $ \star = \star $ has exactly one inhabitant, $ \C_M $ is univalent.

>>>>

\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_univalence}]\label{rem:karoubi-univalent}
  If $ \overline \C $ is univalent, $ \C $ is univalent as well.
\end{lemma}
\begin{proof}
  The fully faithful embedding $ \iota_\C : \C \hookrightarrow \overline \C $ induces equivalences $ (X \cong Y) \simeq (\iota_\C(X) \cong \iota_\C(Y)) $. We also have an equivalence $ (X = Y) \simeq (\iota_\C(X) = \iota_\C(Y)) $, because, as it turns out, any equality between $ X $ and $ Y $ also preserves the identity of $ X $. Therefore, if $ \overline \C $ is univalent, we have a chain of equivalences
  \[ (X = Y) \simeq (\iota_\C(X) = \iota_\C(Y)) \simeq (\iota_\C(X) \cong \iota_\C(Y)) \simeq (X \cong Y), \]
  which shows that $ \C $ is univalent as well.
\end{proof}

\begin{remark}\label{rem:karoubi-not-univalent}
  Note that the converse does not necessarily hold. Consider the commutative monoid consisting of the three matrices
  \[
    a = \begin{pmatrix}
      1 & 0\\0 & 1
    \end{pmatrix}, \quad
    b = \begin{pmatrix}
      1 & 0\\0 & 0
    \end{pmatrix} \quad \text{and} \quad
    c = \begin{pmatrix}
      -1 & 0\\0 & 0
    \end{pmatrix}
  \]
  under matrix multiplication. As we saw in Section \ref{sec:monoid-category}, we can turn this into a category $ \C_M $ with one object $ \star $, and the three morphisms $ a $, $ b $ and $ c $. Only $ a $ is an isomorphism and since $ \star = \star $ has exactly one inhabitant, $ \C_M $ is univalent.

### 2024-11-15 500-previous-work.tex
Add a remark about univalence of L0, and string the remarks about univalence of the category of retracts together

\begin{remark}
  Note that if $ L $ is a nontrivial $ \lambda $-theory, $ \R $ is not a univalent category. To see this, note, for example, that we have an object $ X := \langle \pi_1, \pi_2 \rangle : \R $ (corresponding to the type of `pairs' of $ \lambda $-terms). Since $ L_0 $ is a set, $ X = X $ is a proposition. However, $ X $ has (at least) two automorphisms:
  \[ \langle \pi_1, \pi_2 \rangle \quad \text{and} \quad \langle \pi_2, \pi_1 \rangle. \]
  These are the identity, and the automorphism (of order 2) that swaps the elements of the pair. To see that these are indeed different morphisms, note that applying them (or their lifted versions) to $ (x_{2, 1}, x_{2, 2}) $ gives respectively $ x_{2, 1} $ and $ x_{2, 2} $, which are distinct elements by Lemma \ref{lem:nontrivial-algebraic-theory}.
\end{remark}

>>>>

\begin{remark}
  As we saw in Remarks \ref{rem:karoubi-univalent} and \ref{rem:karoubi-not-univalent}, the Karoubi envelope can only be univalent if the original category is univalent, and even then it is not always univalent. In the case of $ \R $, if $ L $ is a nontrivial $ \lambda $-theory, $ \R $ is not a univalent category. To see this, note, for example, that we have an object $ X := \langle \pi_1, \pi_2 \rangle : \R $ (corresponding to the type of `pairs' of $ \lambda $-terms). Since $ L_0 $ is a set, $ X = X $ is a proposition. However, $ X $ has (at least) two automorphisms:
  \[ \langle \pi_1, \pi_2 \rangle \quad \text{and} \quad \langle \pi_2, \pi_1 \rangle. \]
  These are the identity, and the automorphism (of order 2) that swaps the elements of the pair. To see that these are indeed different morphisms, note that applying them (or their lifted versions) to $ (x_{2, 1}, x_{2, 2}) $ gives respectively $ x_{2, 1} $ and $ x_{2, 2} $, which are distinct elements by Lemma \ref{lem:nontrivial-algebraic-theory}.
\end{remark}

\begin{remark}
  In some cases, we can already conclude that $ \R $ is not univalent because $ L_0 $ is not univalent. For example, if $ L_0 $ has $ \eta $-equality, the single object $ \star : L_0 $ has a nontrivial automorphism given by
  \[ \lambda x_1 x_2 x_3, x_1 x_3 x_2, \]
  so $ L_0 $ is not univalent, so $ \R $ is not univalent.

  On the other hand, if $ L $ is the pure $ \lambda $-calculus with $ \beta $-equality, I believe we can use structural induction to show that $ \star : L_0 $ has no nontrivial automorphisms, so $ L_0 $ is univalent. However, as shown above, $ \R $ is still not univalent.
\end{remark}
